Formal Analysis of Rate-Monotonic Scheduling Implementation via Real-Time Maude

### 2015-10-09 RMS/submission.tex
adding bib file, finishing introduction.

\section{Introduction}

\section{Rate-Monotonic Scheduling Algorithm and the Implementation}

\subsection{Formulation of the Standard Setting}
The frequently used setting of the RMS algorithm for analysis
considers a task set consisting only of $n$ periodic tasks
$\tau_1,\ldots,\tau_n$. Task $\tau_i$ has a period $T_i$ and a
computation time $C_i$. First jobs of all tasks are assumed to be
instantiated simultaneously. This means that jobs corresponding to
task $\tau_i$ are initiated at times $kT_i, k\ge 0$. Deadlines consist
of runability constraints only, that is, the job initiated at time
$kT_i$ has $(k+1)T_i$ as its deadline, which is the initiation time of
the next job. The RMS algorithm chooses the labeling such that $T_1\le
T_2\le \ldots \le T_n$. Consequently, $\tau_i$ receives priority
$i$. The following assumptions are made:

@@

\section{Real-Time Maude}
Real-Time Maude is a language and tool that extends Maude to support
the formal specification and analysis of real-time systems.

@@

\section{Formalizing the RMS Implementation}
Based on the code, we model the implementation under the following
assumptions:

@@

\section{Formal Verification}
Although Real-Time Maude provides many useful commands and tools to
analyze a model, such as, \verb|rewrite| to execute the model,
\verb|search| to search some reachable state(s) satisfying given
properties from the initial state, the Maude Inductive Theorem Prover
(ITP) to interactively prove properties written in \emph{membership
  equational logic}, and so on, we concentrate in this paper on using
the model checking tool to verify the desired properties of our model.

@@

\section{Related Work}

\section{Conclusion}

\end{document}

>>>>

\section{Introduction}
Periodic task scheduling is one of the most studied topics within the
field of real-time systems, due to the large number of control systems
that require cyclic activities. A set of periodic tasks is said to be
\emph{schedulable} with respect to some scheduling algorithm if all
jobs meet their deadlines. \emph{Rate-Monotonic Scheduling}
(\emph{RMS}) algorithm is a \emph{fixed} priority scheduling algorithm
for preemptive hard real-time environments proposed by Liu and
Layland~\cite{DBLP:journals/jacm/LiuL73} in 1973, which assigns
priorities to jobs according to the periods of the corresponding
tasks: the smaller the period, the higher the priority. As it is easy
to implement and is proved to be the \emph{optimal} fixed priority
scheduling algorithm~\cite{DBLP:journals/jacm/LiuL73}, in the sense
that any set of tasks that is schedulable under \emph{some} fixed
priority scheduling algorithm is also schedulable with respect to RMS
algorithm, RMS is widely used in different kinds of real-time
applications, some of which are even safety-critical, such as trains,
avionics and so on.

Liu and Layland also gave a sufficient condition for the
schedulability of a set of $n$ tasks scheduled by RMS algorithm
in~\cite{DBLP:journals/jacm/LiuL73}:
$\displaystyle\Sigma^n_{i=1}C_i/T_i \le n(2^{1/n}-1)$, where $C_i$ and
$T_i$ are the computation time and the period of task $\tau_i$
respectively. Two main directions on RMS have been explored since
then. One is to relax the assumptions on the original RMS model, for
instance, to allow aperiodic
tasks~\cite{DBLP:conf/rtss/LehoczkySS87,DBLP:journals/rts/SpruntSL89,DBLP:conf/rtss/LehoczkyR92,DBLP:journals/tc/StrosniderLS95},
to generalize it to be
\emph{deadline-monotonic}~\cite{DBLP:journals/pe/LeungW82,audsley1993deadline},
to allow resource sharing among tasks~\cite{DBLP:journals/tc/ShaRL90},
to extend it on
multiprocessors~\cite{dhall1978real,DBLP:journals/rts/LopezGDG03,DBLP:journals/tpds/LopezDG04,DBLP:journals/tc/BaruahG03},
to enhance
fault-tolerance~\cite{DBLP:journals/rts/OhS94,DBLP:journals/rts/GhoshMMS98,DBLP:journals/tpds/BertossiMR99}
and so on, making RMS algorithm applicable on more systems. The other
direction is to generate better schedulablity test conditions for the
algorithm and its
extensions~\cite{DBLP:conf/rtss/LehoczkySD89,DBLP:conf/rtss/KuoM91,DBLP:journals/tc/BiniBB03,DBLP:journals/rts/LopezGDG03,DBLP:journals/tc/BaruahG03,gardner1999}. The
RMS algorithm is of such practical importance that we certainly cannot
exhaust all the results in literatures here.

On the other hand, an increasing number of efforts to apply formal
methods, such as model checking and theorem proving, are made to
analyze safety-critical systems for the past few
years~\cite{DBLP:journals/iandc/MeseguerR13,DBLP:journals/cacm/Leroy09,DBLP:conf/sosp/KleinEHACDEEKNSTW09}. However,
as far as we know, few~\cite{DBLP:conf/iceccs/CuiDT14,TianD2011} are
made to analyze the RMS algorithm, none for implementations of RMS.

In this paper, we use \emph{Real-Time Maude}, a \emph{rewriting}-based
modeling language and analyzing tool for real-time systems, to model a
realistic implementation of RMS algorithm and then verify some desired
properties on the built model. The model we built based on the
realistic implementation extends the original ideal one proposed
in~\cite{DBLP:journals/jacm/LiuL73}, by allowing computation time of
tasks to be non-constant, and by considering overheads and other
realistic aspects.

The rest of this paper is organized as follows: Section~\ref{s:rms}
introduces the standard setting of the RMS algorithm proposed
in~\cite{DBLP:journals/jacm/LiuL73} and our considering realistic
implementation. Section~\ref{s:rtmaude} gives a background to
Real-Time Maude. Section~\ref{s:formalism} presents how we model the
RMS implementation using Real-Time Maude. Then
Section~\ref{s:verification} explains how to verify the desired
properties and to analyze the results. Related work is discussed in
Section~\ref{s:relate} before concluding in
Section~\ref{s:conclusion}.

\section{Rate-Monotonic Scheduling Algorithm and the Implementation}
\label{s:rms}
\subsection{Formulation of the Standard Setting}
The frequently used setting of the RMS algorithm for analysis
considers a task set consisting only of $n$ periodic tasks
$\tau_1,\ldots,\tau_n$. Task $\tau_i$ has a period $T_i$ and a
computation time $C_i$. First jobs of all tasks are assumed to be
instantiated simultaneously. This means that jobs corresponding to
task $\tau_i$ are initiated at times $kT_i, k\ge 0$. Deadlines consist
of runability constraints only, that is, the job initiated at time
$kT_i$ has $(k+1)T_i$ as its deadline, which is the initiation time of
the next job. The RMS algorithm chooses the labeling such that $T_1\le
T_2\le \ldots \le T_n$. Consequently, $\tau_i$ receives priority
$i$. The following assumptions are made:

@@

\section{Real-Time Maude}
\label{s:rtmaude}
Real-Time Maude is a language and tool that extends Maude to support
the formal specification and analysis of real-time systems.

@@

\section{Formalizing the RMS Implementation}
\label{s:formalism}
Based on the code, we model the implementation under the following
assumptions:

@@

\section{Formal Verification}
\label{s:verification}
Although Real-Time Maude provides many useful commands and tools to
analyze a model, such as, \verb|rewrite| to execute the model,
\verb|search| to search some reachable state(s) satisfying given
properties from the initial state, the Maude Inductive Theorem Prover
(ITP) to interactively prove properties written in \emph{membership
  equational logic}, and so on, we concentrate in this paper on using
the model checking tool to verify the desired properties of our model.

@@

\section{Related Work}
\label{s:relate}

\section{Conclusion}
\label{s:conclusion}

\bibliographystyle{splncs03}
\bibliography{submission}
\end{document}
